---
title: "BAN 502 Final Project Data Prep"
author: "Mathews, Chris"
date: '2022-06-16'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'asis',     # Can also be set at chunk level
                      comment = NA,
                      prompt  = FALSE,
                      cache   = FALSE,
                      fig.width = 12, 
                      fig.height = 8)
```

## Load the libraries  

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(skimr)
library(mice) #package for imputation
#library(VIM) #visualizing missingness
library(caret)
library(GGally)
library(gridExtra)
library(corrplot)
library(correlationfunnel)
library(forcats)
library(summarytools)
```

## Read the Data  

Read the data and see what we are dealing with.

```{r}
ames <- read_csv("~/BAN502/CourseProject/ames_student.csv")
```


```{r warning=FALSE}
summary(ames)
#stview(summarytools::dfSummary(ames))
```

## Pre-Process Categorical Data

Plenty of character variables. First we need them as factors. Then lets examine the levels to determine whether the variables are balanced or larger a single value.

```{r}
ames = ames %>% mutate_if(is.character,as_factor)
```

Generate a sorted list of factors to use for later transforms

```{r}
sort(names(which(sapply(ames, class) == "factor")))
```

```{r}
factor_list = c("Above_Median", "Alley", "Bldg_Type", "Bsmt_Cond", "Bsmt_Exposure", "Bsmt_Qual", "BsmtFin_Type_1", "BsmtFin_Type_2", "Central_Air", "Condition_1", "Condition_2", "Electrical", "Exter_Cond", "Exter_Qual",    
"Exterior_1st", "Exterior_2nd", "Fence", "Fireplace_Qu", "Foundation", "Functional", "Garage_Cond", "Garage_Finish", "Garage_Qual", "Garage_Type", "Heating", "Heating_QC", "House_Style", "Kitchen_Qual",   
"Land_Contour", "Land_Slope", "Lot_Config", "Lot_Shape", "Mas_Vnr_Type", "Misc_Feature", "MS_SubClass", "MS_Zoning", "Neighborhood", "Overall_Cond", "Overall_Qual", "Paved_Drive", "Pool_QC", "Roof_Matl",
"Roof_Style", "Sale_Condition", "Sale_Type", "Street", "Utilities")
```

Using the factor list above, combine low freqency levels into an "Other" group. Set the value relatively low, but can rerun at higher later on

```{r}
ames <- ames %>%
  mutate(across(factor_list, fct_lump_min,min = 50,other_level = 'Other'))
```

### generate graphics for presentation

```{r message=FALSE, warning=FALSE}
ggplot(ames, aes(x=Heating,fill=Above_Median)) + geom_bar(alpha=0.7) + labs(title="Heating")
ggsave("Heatingfct.png")
```


Seems that cleaning the low frequency levels have some, but a few extremely unbalanced factors remain.

  * Sale_Condtion is defined as the type of sale and includes outliers like foreclosures and family sales. I would expect these types of sales to follow other patterns and should be excluded from this analysis
  * Additionally, the Utilities variable indicates that all properties have all municipal utilities except one. Exclude this property as well.
  * MS_SubClass levels appear to be a combination of columns, remove it
  * Latitude and Longitude could be important in some types of analysis, but proximity to other like properties in likely covered in the Neighborhood variable, drop them.

There are some that are still unbalanced(Street, Alley, Bsmt*), but might prove interesting like the Pool_QC column that indicates whether the property contains a pool. Removed the Pool_area column and will hold in reserve the remaining few categorical variables.  

```{r}
ames <- subset(ames, 
               ames$Sale_Condition == "Normal")
ames <- subset(ames, 
               ames$Utilities == "AllPub")
ames <- select(ames, -c(MS_SubClass, Latitude, Longitude, Utilities, Sale_Condition, Pool_Area))
```

### Finish Categorical Variables Pre-Processing


## Pre-Process Numeric variables

Generate a list of numeric variables

```{r}
sort(names(which(sapply(ames, class) == "numeric")))
```

```{r}
numeric_discrete_list = c("Year_Built", "Year_Remod_Add", "BsmtFin_SF_1", "Bedroom_AbvGr", "Bsmt_Full_Bath", "Bsmt_Half_Bath", "Fireplaces", "Full_Bath", "Garage_Cars", "Half_Bath",  "Kitchen_AbvGr",  "Mo_Sold", "Pool_QC", "TotRms_AbvGrd", "Year_Sold" )
numeric_continous_list = c("Lot_Area", "Lot_Frontage", "Mas_Vnr_Area", "Bsmt_Unf_SF", "BsmtFin_SF_2", "Total_Bsmt_SF", "First_Flr_SF","Second_Flr_SF",  "Low_Qual_Fin_SF", "Gr_Liv_Area", "Enclosed_Porch", "Garage_Area", "Misc_Val", "Open_Porch_SF", "Screen_Porch",  "Three_season_porch", "Wood_Deck_SF" )

```

```{r}
numeric_discrete_list
```

```{r}
p1 = ggplot(ames, aes(x=Garage_Cars,fill=Above_Median)) + geom_bar(alpha=0.7)
p2 = ggplot(ames, aes(x=Half_Bath,fill=Above_Median)) + geom_bar(alpha=0.7)
p3 = ggplot(ames, aes(x=Kitchen_AbvGr,fill=Above_Median)) + geom_bar(alpha=0.7)
p4 = ggplot(ames, aes(x=Mo_Sold,fill=Above_Median)) + geom_bar(alpha=0.7)
p5 = ggplot(ames, aes(x=Pool_QC,fill=Above_Median)) + geom_bar(alpha=0.7)
p6 = ggplot(ames, aes(x=TotRms_AbvGrd,fill=Above_Median)) + geom_bar(alpha=0.7)
p7 = ggplot(ames, aes(x=Year_Sold,fill=Above_Median)) + geom_bar(alpha=0.7)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=2)
```
The above plots indicate that Kitchen_AbvGr and Pool_QC remain unbalanced and are candidates for exclusion

```{r}
ames <- select(ames, -c(Kitchen_AbvGr, Pool_QC))
```

```{r}
p1 = ggplot(ames, aes(x=Year_Built,fill=Above_Median)) + geom_bar(alpha=0.7)
p2 = ggplot(ames, aes(x=Year_Remod_Add,fill=Above_Median)) + geom_bar(alpha=0.7)
p3 = ggplot(ames, aes(x=BsmtFin_SF_1,fill=Above_Median)) + geom_bar(alpha=0.7)
p4 = ggplot(ames, aes(x=Bedroom_AbvGr,fill=Above_Median)) + geom_bar(alpha=0.7)
p5 = ggplot(ames, aes(x=Bsmt_Full_Bath,fill=Above_Median)) + geom_bar(alpha=0.7)
p6 = ggplot(ames, aes(x=Bsmt_Half_Bath,fill=Above_Median)) + geom_bar(alpha=0.7)
p7 = ggplot(ames, aes(x=Fireplaces,fill=Above_Median)) + geom_bar(alpha=0.7)
p8 = ggplot(ames, aes(x=Full_Bath,fill=Above_Median)) + geom_bar(alpha=0.7)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8,  ncol=2)
```

### Generate graphics for presentation

```{r}
p1 =ggplot(ames, aes(x=Bsmt_Half_Bath,fill=Above_Median)) + geom_bar(alpha=0.7) + labs(title = "Basement Half Baths")
ggsave("BsmtHalfBath.png")
p2 = ggplot(ames, aes(x=TotRms_AbvGrd,fill=Above_Median)) + geom_bar(alpha=0.7) + labs(title = "Total Rooms Above Ground")
ggsave("TotRms_AbvGrd.png")
grid.arrange(p1, p2, ncol=2)
```


Bsmt_Half_Bath might need to be excluded.

```{r}
ames <- select(ames, -c(Bsmt_Half_Bath))
```

```{r}
numeric_continous_list
```

```{r}
p2 = ggplot(ames, aes(x = Lot_Area)) +  
  geom_histogram(binwidth = 2000, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
p4 = ggplot(ames, aes(x = Lot_Frontage)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
p6 = ggplot(ames, aes(x = Mas_Vnr_Area)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
p8 = ggplot(ames, aes(x = Bsmt_Unf_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
grid.arrange(p2, p4, p6, p8, ncol = 2)
```

### Generate graphics for presentation

```{r}
p1 = ggplot(ames, aes(x = Lot_Area)) +  
  geom_histogram(binwidth = 2000, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal()
ggsave("Lot_Area.png")
p2 = ggplot(ames, aes(x = Lot_Frontage)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
ggsave("Lot_front.png")
grid.arrange(p1, p2, ncol=2)
```


Mas_Vnr_Area is the would likely indicate no Masonary veneer and the Bsmt_Unf_Sf might indicate no basement.
The Lot_Area is of concern and we like need to exclude the largest lots. The median value is 9525 sq ft, there are 11 properties with lot area greater than 1 acre(43560 sq ft). We will be excluding them from the analysis.
Additionally, it is unlikely that properties sold have zero frontage and is likely keying error. We will impute the median value based on Lot_Shape and replace the zero's.

```{r}
ames <- ames %>% filter(Lot_Area <= 43560)
```

```{r}
sprintf("For: \n") %>% cat()
for (i in unique(ames$Lot_Shape))
    {
    ames_i <- ames %>% filter(Lot_Shape==i)
    sprintf(
            fmt = " -%s houses, the mean Lot_Frontage is %.2f and median Lot_Frontage is %.2f\n",
            i, mean(ames_i$Lot_Frontage, na.rm=TRUE), median(ames_i$Lot_Frontage, na.rm=TRUE)
            ) %>% cat()
}
```

```{r}
ames <- ames %>% 
  group_by(Lot_Shape) %>% 
  mutate(Lot_Frontage = ifelse(Lot_Frontage == 0, 
                            median(Lot_Frontage, na.rm = TRUE), 
                            Lot_Frontage))
```

```{r}
p2 = ggplot(ames, aes(x = Lot_Area)) +  
  geom_histogram(binwidth = 2000, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
p4 = ggplot(ames, aes(x = Lot_Frontage)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
p6 = ggplot(ames, aes(x = Mas_Vnr_Area)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
p8 = ggplot(ames, aes(x = Bsmt_Unf_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
grid.arrange(p2, p4, p6, p8, ncol = 2)
```
### Generate a graphic for presentation

```{r}
ggplot(ames, aes(x = Mas_Vnr_Area)) +  
  geom_histogram(binwidth = 50, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 2000, by = 40)) + 
  theme_minimal()
ggsave("Mas_Vnr_Area.png")
```


Mas_Vnr_Area and Bsmt_Unf_Sf should be converted to binary values indicating whether Masonary Veneer and Basement is present for the property

```{r}
ames$Mas_Vnr_Present <- factor(ifelse(ames$Mas_Vnr_Area== 0,"No","Yes"))
ggplot(ames, aes(x=Mas_Vnr_Present,fill=Above_Median)) + geom_bar(alpha=0.7)
ames <- select(ames, -c(Mas_Vnr_Area))
```

```{r}
ames$Bsmt_Present <- factor(ifelse(ames$Bsmt_Unf_SF== 0,"No","Yes"))
ggplot(ames, aes(x=Bsmt_Present,fill=Above_Median)) + geom_bar(alpha=0.7)
ames <- select(ames, -c(Bsmt_Unf_SF))
```

```{r}
numeric_continous_list
```

```{r}
p2 = ggplot(ames, aes(x = Wood_Deck_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
grid.arrange(p2, ncol = 2)
```

Likely need to convert to binary Yes for Wood Deck present and No for No Deck.

```{r}
ames$Wood_Deck <- factor(ifelse(ames$Wood_Deck_SF== 0,"No","Yes"))
ggplot(ames, aes(x=Wood_Deck,fill=Above_Median)) + geom_bar(alpha=0.7)
ames <- select(ames, -c(Wood_Deck_SF))
```

```{r}
p2 = ggplot(ames, aes(x = Misc_Val)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
p4 = ggplot(ames, aes(x = Open_Porch_SF)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
p6 = ggplot(ames, aes(x = Screen_Porch)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
p8 = ggplot(ames, aes(x = Three_season_porch)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
grid.arrange(p2, p4, p6, p8, ncol = 2)
```

All the above variables can likley be excluded or converted to binary state.

```{r}
ames$Misc_Val_fct <- factor(ifelse(ames$Misc_Val== 0,"No","Yes"))
ames$Open_Porch_fct <- factor(ifelse(ames$Open_Porch_SF== 0,"No","Yes"))
ames$Screen_Porch_fct <- factor(ifelse(ames$Screen_Porch== 0,"No","Yes"))
ames$Three_season_porch_fct <- factor(ifelse(ames$Three_season_porch== 0,"No","Yes"))

p1 = ggplot(ames, aes(x=Misc_Val_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
p2 = ggplot(ames, aes(x=Open_Porch_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
p3 = ggplot(ames, aes(x=Screen_Porch_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
p4 = ggplot(ames, aes(x=Three_season_porch_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
grid.arrange(p1, p2, p3, p4, ncol=2)

ames <- select(ames, -c(Misc_Val, Open_Porch_SF, Screen_Porch, Three_season_porch))
```
Even converting to binary factor for all but Open Porch appears lopsided

```{r}
p2 = ggplot(ames, aes(x = Low_Qual_Fin_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
p4 = ggplot(ames, aes(x = Gr_Liv_Area)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
p6 = ggplot(ames, aes(x = Enclosed_Porch)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
p8 = ggplot(ames, aes(x = Garage_Area)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
grid.arrange(p2, p4, p6, p8, ncol = 2)
```

Low_Qual_Fin_SF and Enclosed_Porch are likely candidates for exclusion or conversion to binary variables

```{r}
ames$Low_Qual_Fin_fct <- factor(ifelse(ames$Low_Qual_Fin_SF== 0,"No","Yes"))
ames$Enclosed_Porch_fct <- factor(ifelse(ames$Enclosed_Porch== 0,"No","Yes"))

p1 = ggplot(ames, aes(x=Misc_Val_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
p2 = ggplot(ames, aes(x=Open_Porch_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
grid.arrange(p1, p2, ncol=2)

ames <- select(ames, -c(Low_Qual_Fin_SF, Enclosed_Porch))
```
Misc_Val_fct can still likely be removed.

```{r}
p2 = ggplot(ames, aes(x = BsmtFin_SF_2)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 200)) + 
  theme_minimal() 
p4 = ggplot(ames, aes(x = Total_Bsmt_SF)) +  
  geom_histogram(binwidth = 25, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 25)) + 
  theme_minimal() 
p6 = ggplot(ames, aes(x = First_Flr_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
p8 = ggplot(ames, aes(x = Second_Flr_SF)) +  
  geom_histogram(binwidth = 20, fill = "lightsalmon2", color = "black") +
  scale_x_continuous(labels = comma, breaks = seq(0, 3500, by = 20)) + 
  theme_minimal() 
grid.arrange(p2, p4, p6, p8, ncol = 2)
```
BsmtFin_SF_2 and Second_Flr_SF are both like candidates for exclusion.

```{r}
ames$BsmtFin_fct <- factor(ifelse(ames$BsmtFin_SF_2== 0,"No","Yes"))
ames$Second_Flr_fct <- factor(ifelse(ames$Second_Flr_SF== 0,"No","Yes"))

p1 = ggplot(ames, aes(x=BsmtFin_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
p2 = ggplot(ames, aes(x=Second_Flr_fct,fill=Above_Median)) + geom_bar(alpha=0.7)
grid.arrange(p1, p2, ncol=2)

ames <- select(ames, -c(BsmtFin_SF_2, Second_Flr_SF))
```

```{r}
ggplot(ames, aes(x = Year_Sold, fill = Above_Median)) +
  geom_bar(position = "dodge", alpha=0.65) +
  labs(title="Counts of Homes Sold By Year")
ggsave("CountByYearSold.png")
```

```{r}
ames_binarized <- ames %>%
    binarize(n_bins = 4, thresh_infreq = 0.01, name_infreq = "Other_Bin", one_hot=TRUE)
sort(colnames(ames_binarized))
```

```{r fig.width=12, fig.height=16}
ames_correlated <- ames_binarized %>%
    correlate(target = Above_Median__Yes)
ames_correlated %>%
    plot_correlation_funnel(interactive=FALSE)
```
```{r warning=FALSE}
ames_final <- ames %>% select(c(Above_Median, Full_Bath, Kitchen_Qual, Bsmt_Qual, Heating_QC, Exter_Qual, Foundation, Fireplaces, Garage_Cars,  Open_Porch_fct, BsmtFin_Type_1, Gr_Liv_Area, MS_Zoning, Lot_Shape, Lot_Frontage, Lot_Area, Year_Remod_Add, Year_Built ))
saveRDS(ames_final, "./AMESFINAL.RDS")
write.csv(ames_final, "./amesfinal.csv")


summary(ames_final)
#stview(dfSummary(ames_final))
```

