---
title: "Decision Tree"
author: "Mathews, Chris"
date: '2022-06-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)
library(summarytools)
library(rpart) #for classification trees
library(rpart.plot) #for plotting trees
library(RColorBrewer) #better visualization of classification trees
library(rattle) #better visualization of classification trees
library(caret)
```

```{r}
ames <- readRDS("AMESFINAL.RDS")
#ames <- ames %>% filter(Kitchen_Qual != "Other") %>% filter(Bsmt_Qual != "Other")
```


Now we'll split the data.  
```{r}
set.seed(123) 
ames_split = initial_split(ames, prop = 0.7, strata = Above_Median) #70% in training
train = training(ames_split) 
test = testing(ames_split)
```

Let's build a classification tree.  
```{r}
ames_recipe = recipe(Above_Median ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

ames_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(ames_recipe)

ames_fit = fit(ames_wflow, train)
```

```{r}
tree = ames_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree) 
```
```{r}
tree$variable.importance
vip(tree)
```




Look at the "rpart" complexity parameter "cp".    
```{r}
ames_fit$fit$fit$fit$cptable
```

Create our folds  
```{r}
set.seed(234)
folds = vfold_cv(train, v = 5)
```


```{r}
ames_recipe = recipe(Above_Median ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                          levels = 25) #try 25 sensible values for cp

ames_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(ames_recipe)

tree_res = 
  ames_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res
```

```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

```{r}
final_wf = 
  ames_wflow %>% 
  finalize_workflow(best_tree)
```

```{r}
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree, tweak = 1.5) 

```
```{r}
vip(tree)
saveRDS(tree, "./dt_ames_final_tree.rds")
```



Predictions on training set  
```{r}
treepred = predict(final_fit, train, type = "class")
head(treepred)
```

Caret confusion matrix and accuracy, etc. calcs  
```{r}
confusionMatrix(treepred$.pred_class,train$Above_Median,positive="Yes") #predictions first then actual
```

Predictions on testing set  
```{r}
treepred_test = predict(final_fit, test, type = "class")
head(treepred_test)
```

Caret confusion matrix and accuracy, etc. calcs  
```{r}
confusionMatrix(treepred_test$.pred_class,test$Above_Median,positive="Yes") #predictions first then actual
```

```{r}
ames_recipe = recipe(Above_Median ~., train) %>% 
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% #don't forget the model = TRUE flag
  set_mode("classification")

tree_grid = expand.grid(cost_complexity = seq(0.001,0.01,by=0.001))

ames_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(ames_recipe)

tree_res = 
  ames_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res
```

```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

```{r}
final_wf = 
  ames_wflow %>% 
  finalize_workflow(best_tree)
```

```{r}
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree, tweak = 1.5) 

```
Predictions on training set  
```{r}
treepred = predict(final_fit, train, type = "class")
head(treepred)
```

Caret confusion matrix and accuracy, etc. calcs  
```{r}
confusionMatrix(treepred$.pred_class,train$Above_Median,positive="Yes") #predictions first then actual
```

Predictions on testing set  
```{r}
treepred_test = predict(final_fit, test, type = "class")
head(treepred_test)
```

Caret confusion matrix and accuracy, etc. calcs  
```{r}
confusionMatrix(treepred_test$.pred_class,test$Above_Median,positive="Yes") #predictions first then actual
```

```{r}
saveRDS(final_fit, "./classtree_ames.rds")
```

```{r}
library(vip)
vip(final_fit$fit)
```

